---
title: "Assignment 6 Markdown"
author: "Autumn Hodgins"
date: "2025-02-10"
output: html_document
---
Markdown for Assignment 6

Loading all the required packages:
```{r}
library(doParallel)
library(parallel)
library(foreach)
```
Checking how many cores my computer has.
```{r}
detectCores() 
```
My computer has 8 cores.


Generating a series of 100,000 numbers, calculating the mean, and repeating this 4000 times. The output is the run time that it takes for my computer to complete this program.
```{r}
start <- Sys.time()

for(i in 1:4000){
  random_numbers <- rnorm(100000, mean = 10, sd = 3)
  mean <- mean(random_numbers)
}

end <- Sys.time()

time_to_run <- end - start
print(time_to_run)
```
I predict that this would take 7 times less with a multi-threaded program, as I will be using 7 cores (leaving 1 core for other processing in my computer). Here is the expected time:

```{r}
print(time_to_run/7)
```
Running this as a parallel program, and outputting the run time:
```{r}
start <- Sys.time()

Cores <- parallel::makeCluster(detectCores()-1)
doParallel::registerDoParallel(Cores)

foreach(i=1:4000, .combine=rbind) %dopar%{
  random_numbers <- rnorm(100000, mean = 10, sd = 3)
  mean <- mean(random_numbers)  
}

parallel::stopCluster(Cores)

end <- Sys.time()

time_to_run <- end - start
print(time_to_run)
```
The time this program took to run was longer than the expected value. This is because my estimate assumed that the run time could be evenly split among 7 cores, and did not take into account the run time that could not be shortened by running in parallel, also known as the overhead.


